Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-08-11 writeups/dissertation/dissertation.tex
foo

\title{Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow}
\shorttitle{Comparing memory models of tag retrieval}
%\shorttitle{}

@@

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
  \caption{Best-fit decay rate}
  \label{figPriorDecay}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
  \caption{Model accuracy}
  \label{figPriorAcc}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
  \caption{Impact of adding model offset for StackOverflow}
  \label{figContextOffsetSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
  \caption{Impact of adding model offset for Twitter}
  \label{figContextOffsetT}
\end{figure}

@@

% FIXME: Cite is too recent
The same previously-derived stoplist from \textcite{Salton2014} that \textcite{Sahlgren2008} used was used for this analysis,
The same entropy weighting method using by \textcite{Stanley2013} was examined,
and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.

@@

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
  \label{figContextPriorneededSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
  \caption{Model accuracy for various word order methods for the random permutation model for Twitter}
  \label{figContextPriorneededT}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
  \caption{Unique variance in prior component for popular StackOverflow users}
  \label{figPUAddPrior}
\end{figure}

@@

\subsubsection{Result}

@@

@@

Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
So it is most likely the case that the task is not done solely with declarative memory processes.

Or, perhaps authors are taking advantage of context information that is not currently represented in the model.
For example, time of day might be a useful predictor of tag use.
\textcite{Cohn2012} built a tag-prediction model that was based primarily on temporal cycles of tag use.
This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
This would allow time of day information to be used alongside the rest of the context information in the words of the post.
A similar technique could be used to add additional context predictors to the model.
These co-occurrences could either be added on as additional columns to the existing co-occurrence matrix,
or a separate co-occurrence matrix could be used for each additional context component, similar to how separate matrices and context predictors were created for the words in the title and body of a post. 
Including additional context predictors such as time of day may help improve model accuracy,
and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
That is, perhaps this tagging process is completely declarative, and the remaining cues that the author is using are yet to be discovered.

>>>>

\title{Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow}
\shorttitle{Comparing memory models of tag retrieval}
%\shorttitle{}

@@

\def\committee{
  Michael D. Byrne, Chair \\
  Professor, Psychology and Computer Science \and
  Phillip Kortum \\
  Assistant Professor, Psychology \and
  David Lane \\
  Associate Professor, Psychology, Statistics, and Management \and
  Devika Subramanian \\
  Professor, Computer Science and Electrical Engineering \\
}

% Replace the title page in the apa6 style sheet with the title page from the Rice template
% The title page style was taken from Rice University LaTeX PhD template: http://scholarship.rice.edu/handle/1911/21747
% The following code for the abstract page was taken from the apa6 \maketitle command.
\makeatletter
\def\makeDisCover{
  \def\maketitle{
    \begin{titlepage}
      \singlespacing
      \null
      \vfill
      \vbox to 0pt{\vss\vbox to 8.75in{\parskip 0pt \parindent 0pt\centering
	{\uppercase\expandafter{Rice University} \par} \vskip 0pt plus 3fil
	{\bf \@title \par} \vskip 0pt plus 1.3fil
	{by \par} \vskip 0pt plus 1fil
	{\bf \@author \par} \vskip 0pt plus 3fil
	{\sc A Thesis Submitted \\
	  in Partial Fulfillment of the \\
	Requirements for the Degree \par} \vskip 0pt plus 1.1fil
	{\bf Doctor of Philosophy \par} \vskip 0pt plus 4.5fil
	\begingroup
	\leftskip .45\textwidth \rightskip 0pt plus 1fil
	{\sc Approved, Thesis Committee: \par} \vskip 2.5em
	\moveright\leftskip\vbox{\hrule width \leftskip}
	\def\and{\par \vskip 2.25em minus 1em\moveright\leftskip\vbox{\hrule width \leftskip}}
	\committee
	\par \endgroup
	\vskip 0pt plus 2fil
	{Houston, TEXAS \par} \vskip 0pt 
	{\monthnameenglish{} \number\year \par}
      }}
      \doublespacing
      \newpage
  %BDB\hyphenpenalty 10000
      \fussy
      \@ifundefined{@abstract}{}{%
	\section{\normalfont\normalsize\abstractname}% BDB
	\noindent\@abstract\par% BDB
	\@ifundefined{@keywords}{}{%
	  \setlength{\parindent}{0.4in}% BDB
	  \indent\textit{\keywordname:} \@keywords%
	}%
	\newpage
      }
      \@ifundefined{def@donotrepeattitle}{
	\section{\protect\normalfont{\@title}}
      }{}%
      \raggedright%
      \setlength{\parindent}{0.4in}%
    \end{titlepage}}
  }
\makeatother

%\makeDisCover

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-d--crop.pdf}}}
  \caption{Overall best-fit decay rate for StackOverflow and Twitter}
  \label{figPriorDecay}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc--crop.pdf}}}
  \caption{Overall model accuracy for StackOverflow and Twitter}
  \label{figPriorAcc}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetSO-crop.pdf}}}
  \caption{Impact of adding offset component for StackOverflow}
  \label{figContextOffsetSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-OffsetT-crop.pdf}}}
  \caption{Impact of adding offset component for Twitter}
  \label{figContextOffsetT}
\end{figure}

@@

The same 571-word Cornell SMART stoplist used by \textcite{Sahlgren2008} was also used for this analysis,
The same entropy weighting method using by \textcite{Stanley2013} was examined,
and the same method of removing the high-frequency words from \citeauthor{Sahlgren2008} was used.

@@

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddSO-crop.pdf}}}
  \caption{Model accuracy when prior is removed for StackOverflow randomly-sampled dataset}
  \label{figContextPriorneededSO}
\end{figure}

\begin{figure}[!htbp]
  \scalebox{.6}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddT-crop.pdf}}}
  \caption{Model accuracy when prior is removed for Twitter popular-hashtags dataset}
  \label{figContextPriorneededT}
\end{figure}

@@

\begin{figure}[!htbp]
  \scalebox{.8}{\resizebox{\linewidth}{!}{\includegraphics{compareMeanDV-acc-priorAddPUserSO-crop.pdf}}}
  \caption{Model accuracy when prior is removed for StackOverflow popular-users dataset}
  \label{figPUAddPrior}
\end{figure}

@@

@@

\subsection{General Results}

Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.

\begin{table}[!ht]
  \caption{Summary of accuracy for each model and dataset}
  \label{tabOverallResults}
  {\tabulinesep=1.2mm
    \begin{tabu}{llll}
      \hline
	Site 		& Dataset		& Model						& Accuracy \\
	\hline
	StackOverflow	& popular-users		& Bayesian with entropy				& \num{.63} \\
			& 			& RP with entropy and log odds 			& \num{.58} \\
			& randomly-sampled	& Bayesian with entropy 			& \num{.63} \\
			& 			& RP with entropy and log odds			& \num{.63} \\
	Twitter		& popular-users		& Bayesian with entropy and user sji		& \num{.39} \\
			& 			& RP with entropy, log odds, and user sji	& \num{.39} \\
			& popular-hashtags	& Bayesian with entropy				& \num{.32} \\
			& 			& RP with entropy and log odds			& \num{.25} \\
	\hline
    \end{tabu}
  }
\end{table}

After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
Both models performed nearly equal for the StackOverflow datasets tested,
and both models were much more accurate when tested on StackOverflow compared to Twitter.

@@

Overall, both the random permutation and Bayesian declarative memory models fit the observed tagging behavior in StackOverflow and Twitter relatively well.
The fits for StackOverflow in particular are quite high, especially given the large tag space on the site.
These results provide support that choosing a hashtag when composing a tweet and tagging a StackOverflow post is a long-term memory retrieval process.
However, both models were not 100 percent accurate, and the amount of variance left to predict was much larger than what would be observed if retrieval noise was included in the model.
So it is most likely the case that the task is not done solely with declarative memory processes.

Or, perhaps authors are taking advantage of context information that is not currently represented in the model.
For example, time of day might be a useful predictor of tag use.
\textcite{Cohn2012} built a tag-prediction model that was based primarily on temporal cycles of tag use.
This technique could be naturally incorporated into a co-occurrence framework by including time of day as an additional context by tag co-occurrence for each post.
This would allow time of day information to be used alongside the rest of the context information in the words of the post.
A similar technique could be used to add additional context predictors to the model.
These co-occurrences could either be added on as additional columns to the existing co-occurrence matrix,
or a separate co-occurrence matrix could be used for each additional context component, similar to how separate matrices and context predictors were created for the words in the title and body of a post. 
Including additional context predictors such as time of day may help improve model accuracy,
and perhaps it is the case that the rest of the variance left to predict is from these context components that the author is using but are not yet included in the model.
